ESBMC-GPU \\ A Context-Bounded Model Checking Tool \\ to Verify CUDA Programs

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
minor changes
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

\vspace{-1mm}

@@

\vspace{-3mm}

@@

\vspace{-2mm}

@@

ESBMC was designed to handle multi-threaded software, through the use of an API called Portable Operating System Interface (POSIX -- ISO/IEC 9945)~\cite{posix:2008}. Thus, ESBMC-GPU applies a combination of processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths. Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).
%Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.

\noindent \textbf{2. Two-threads Analysis.} Similarly to GPUVerify~\cite{betts:2012} and PUG~\cite{Li:2010}, ESBMC-GPU also reduces the number of threads (to only two elements), during the verification of CUDA programs, by considering a NVIDIA Fermi GPU architecture, in order to improve verification time and avoid the state-space explosion problem. In CUDA programs, whilst threads execute the same parametrized kernel, only two of them are necessary for conflict check. Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation~\cite{pereira:2016}.

@@

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also \textcolor{blue}{reduce the state space}, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.% \textcolor{red}{similares? entao, qual eh o impacto de se verificar apenas um? eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.

@@

\vspace{-4mm}

@@

\vspace{-1mm}

In this part, ESBMC-GPU usage is demonstrated, by using the CUDA program shown in Fig.~\ref{fig:illustrative-example}. First of all, users must replace the default kernel call (line $16$) by an intrinsic function of ESBMC-GPU (line $17$). Then, the resulting CUDA program can be passed to the command-line version of ESBMC-GPU, as follows: \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>} {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},  where {\tt $<$file$>$.cu} is the CUDA program, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-} {\tt CUDA-OM$>$} is the location of the COM library.

%\begin{center}
%\vspace{-2 mm}
%\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}

%\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
%\vspace{-2 mm}
%\end{center}

@@

\vspace{-7mm}

@@

 In order to answer both questions, all benchmarks were verified with $4$ GPU verifiers (ESBMC-GPU $v2.0$, GKLEE $v2012$, GPUVerify $v1811$, and CIVL $v1.7.1$), on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, with $16$ GB of RAM, running Ubuntu 14.04 OS. Importantly, all presented execution times are actually CPU times, {\it i.e.}, only the elapsed time periods spent in the allocated CPUs, which was measured with the {\tt times} system call (POSIX system). An overview of the experimental results is shown in Fig.~\ref{figure:experimental}, where {\it True} represents bug-free benchmarks, {\it False} represents buggy benchmarks, {\it Not supported} represents benchmarks that could not be verified, {\it Correct} represents the percentage of benchmarks correctly verified, and {\it Incorrect} represents the percentage of benchmarks incorrectly verified ({\it i.e.}, a verification tool reports an unexpected result). As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}. More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%). %\textcolor{red}{e os outros? quais sao as vantagens? esse paragrafo ficou incompleto} \textcolor{blue}{estamos focando apenas nos aspectos do ESBMC porque nao temos espaco para discutir sobre os demais, isso estara na pagina online} \textcolor{green}{Concordo em colocar o texto explicativo na pagina online e referenciar.}

{\bf Limitations.} ESBMC-GPU was unable to correctly verify $24$ benchmarks, which are related to constant memory access ($2$\%), CUDA's specific libraries ($4.5$\%), and the use of pointers to functions, structures, and {\tt char} type variables, when passed as kernel call arguments ($4.5$\%). In addition, it only reported $3$\% of incorrect true and $1$\% of incorrect false results.

%In addition, ESBMC-GPU had $23$ benchmarks that were not supported. These are related to constant memory access ($3$), the use of CUDA's specific libraries (\textit{e.g., curand.h}) ($7$), and the use of pointers to functions, structures, and char type variables used as kernel call arguments ($13$).

@@

\vspace{-2mm}

\vspace{-8mm}

>>>>

@@

%\vspace{-3mm}

@@

%\vspace{-2mm}

@@

ESBMC was designed to handle multi-threaded software, through the use of an API called Portable Operating System Interface (POSIX -- ISO/IEC 9945)~\cite{posix:2008}. Thus, ESBMC-GPU applies a combination of processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths. Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).\\
%Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.

\noindent \textbf{2. Two-threads Analysis.} Similarly to GPUVerify~\cite{betts:2012} and PUG~\cite{Li:2010}, ESBMC-GPU also reduces the number of threads (to only two elements), during the verification of CUDA programs, by considering a NVIDIA Fermi GPU architecture, in order to improve verification time and avoid the state-space explosion problem. In CUDA programs, whilst threads execute the same parametrized kernel, only two of them are necessary for conflict check. Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation~\cite{pereira:2016}.\\

@@

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also \textcolor{blue}{reduce the state space}, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.\\% \textcolor{red}{similares? entao, qual eh o impacto de se verificar apenas um? eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.

@@

%\vspace{-4mm}

@@

In this part, ESBMC-GPU usage is demonstrated, by using the CUDA program shown in Fig.~\ref{fig:illustrative-example}. First of all, users must replace the default kernel call (line $16$) by an intrinsic function of ESBMC-GPU (line $17$). Then, the resulting CUDA program can be passed to the command-line version of ESBMC-GPU, as follows: %\texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>} {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},

\begin{center}
%\vspace{-2 mm}
\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}

\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
%\vspace{-2 mm}
\end{center}

where {\tt $<$file$>$.cu} is the CUDA program, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-} {\tt CUDA-OM$>$} is the location of the COM library.

@@

%\vspace{-7mm}

@@

 In order to answer both questions, all benchmarks were verified with $4$ GPU verifiers (ESBMC-GPU $v2.0$, GKLEE $v2012$, GPUVerify $v1811$, and CIVL $v1.7.1$), on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, with $16$ GB of RAM, running Ubuntu 14.04 OS. Importantly, all presented execution times are actually CPU times, {\it i.e.}, only the elapsed time periods spent in the allocated CPUs, which was measured with the {\tt times} system call (POSIX system). An overview of the experimental results is shown in Fig.~\ref{figure:experimental}, where {\it True} represents bug-free benchmarks, {\it False} represents buggy benchmarks, {\it Not supported} represents benchmarks that could not be verified, {\it Correct} represents the percentage of benchmarks correctly verified, and {\it Incorrect} represents the percentage of benchmarks incorrectly verified ({\it i.e.}, a verification tool reports an unexpected result). As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}. More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).\\ %\textcolor{red}{e os outros? quais sao as vantagens? esse paragrafo ficou incompleto} \textcolor{blue}{estamos focando apenas nos aspectos do ESBMC porque nao temos espaco para discutir sobre os demais, isso estara na pagina online} \textcolor{green}{Concordo em colocar o texto explicativo na pagina online e referenciar.}

{\bf Limitations.} ESBMC-GPU was unable to correctly verify $24$ benchmarks, which are related to constant memory access ($2$\%), CUDA's specific libraries ($4.5$\%), and the use of pointers to functions, structures, and {\tt char} type variables, when passed as kernel call arguments ($4.5$\%). In addition, it only reported $3$\% of incorrect true and $1$\% of incorrect false results.\\

@@
